---
params:
  run: x
  compartment: x
  residuals: x
  keep_time_zero: x
  pvalue: x
  boxplot_coefficient: x
header-includes:
  - \usepackage{fancyhdr}
  - \usepackage{lipsum}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[L]{Run ID{{:}} `r str_replace_all(params$run$info$run_id, '_', '\\\\_')`}
  - \renewcommand\footrulewidth{1pt}
  - \fancyfoot[L]{`r str_replace_all(params$run$info$path, '_', '\\\\_')`}
  - \fancyfoot[R]{\thepage}
  - \fancypagestyle{plain}{\pagestyle{fancy}}
title: "Outliers detection"
subtitle:  "Based on `r params$residuals`"
date: "`r paste('Report generation:', format(Sys.time(), '%m/%d/%Y (%H:%M)'))`"
output: 
  pdf_document: default
  html_vignette:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

run <- params$run
```

## Info

```{r}
filters <- attr(run$tables$pmxploitab, "filters")

has_filters <- !is.null(filters)
n_obs <- run$info$number_of_observations
n_filtered_rows <- nrow(run$tables$pmxploitab)

filter_text <- ifelse(has_filters & (n_filtered_rows != n_obs),
                      sprintf("%s (filtered from %s)", n_filtered_rows, n_obs),
                      n_obs)
```

Number of observations: `r filter_text`

`r if(has_filters) {"Filters:"}`
```{r}
if(has_filters)
  knitr::kable(tibble(condition = as.character(filters)), align = "c")
```

# Normality check

```{r}
cleaned_df <- run %>%
  filter(!is.na(DV), MDV == 0, CMT == params$compartment) %>% 
  pluck("tables") %>% 
  pluck("pmxploitab")

if(!params$keep_time_zero)
  cleaned_df <- cleaned_df %>% filter(TIME > 0)
```

## Q-Q plot
```{r}
ggplot(cleaned_df)+
    geom_qq(aes_string(sample = params$residuals))+
    labs(x = "Normal distribution quantiles", y = paste(params$residuals, "distribution quantiles"))
```

## Kolmogorov-Smirnov test
```{r}
values <- cleaned_df[[params$residuals]]

ks <- ks.test(unique(values), "pnorm", mean = mean(values), sd = sd(values))

stats_df <- data_frame(mean = mean(values), sd = sd(values)) %>%
  rename(!!!setNames(nm = sprintf("%s (%s)", params$residuals, colnames(.)), colnames(.)))

test_df <- broom::tidy(ks) %>%
  mutate(p.value = format(p.value, scientific = TRUE))# %>% 

ks_df <- bind_cols(stats_df, test_df)

knitr::kable(ks_df, align = "c")
```

```{r echo = FALSE}
normality_ok <- ks$p.value > params$pvalue
normality_msg <- (if(normality_ok){
  sprintf("Normality hypothesis cannot be rejected (p-value > %s)", params$pvalue)
} else {
  sprintf("Normality hypothesis is rejected (p-value < %s)", params$pvalue)
})
```

`r normality_msg`

Note: duplicate values are considered as one unique value

# Grubb's test (T-procedure)

```{r}
grubbs <- run %>% detect_outliers(
  compartment = params$compartment,
  residuals = params$residuals,
  method = "grubbs",
  grubbs_pvalue_threshold = params$pvalue, 
  keep_time_zero = params$keep_time_zero)

n_outliers <- nrow(grubbs$outliers)
n_source <- nrow(grubbs$source)

grubbs_msg <- (
  if(n_outliers == 0) {
    sprintf("No outliers detected in %s observations.", n_source)
  } else if (n_outliers == 1) {
    sprintf("One outlier detected in %s observations (%s).", n_source, scales::percent(n_outliers / n_source))
  } else {
    sprintf("%s outliers detected in %s observations (%s).", n_outliers, n_source, scales::percent(n_outliers / n_source))
  })
```

`r grubbs_msg`

## Outliers

```{r}
if(nrow(grubbs$outliers) > 0) 
  knitr::kable(grubbs$outliers, align = "c")
```


Note: Grubb's test makes the assumption that data is normally distributed.

# Non-parametric

```{r, echo = FALSE}
boxplot <- run %>% 
  detect_outliers(compartment = params$compartment,
                  residuals = params$residuals,
                  method = "boxplot",
                  boxplot_coefficient = params$boxplot_coefficient, 
                  keep_time_zero = params$keep_time_zero)

n_outliers <- nrow(boxplot$outliers)
n_source <- nrow(boxplot$source)

boxplot_msg <- (
  if(n_outliers == 0) {
    sprintf("No outliers detected in %s observations.", n_source)
  } else if (n_outliers == 1) {
    sprintf("One outlier detected in %s observations (%s).", n_source, scales::percent(n_outliers / n_source))
  } else {
    sprintf("%s outliers detected in %s observations (%s).", n_outliers, n_source, scales::percent(n_outliers / n_source))
  })
```

`r boxplot_msg`

Observations are detected as outliers if they are outside of the following interval:

$$\mathopen{[}F_f - k \cdot IQR\,;F_t + k \cdot IQR\mathclose{]}$$

with `Ff` being the first quartile, `Ft` the third quartile, `IQR` the inter-quartile range `(Ft-Ff)` and `k` is a coefficient (length of the whiskers as multiple of `IQR`; here `k = ``r params$boxplot_coefficient`).

## Outliers

```{r}
if(nrow(boxplot$outliers) > 0)
  knitr::kable(boxplot$outliers, align = "c")
```

## Box-and-whisker plot

```{r}
g <- ggplot(mapping = aes_string(x = factor(1), y = boxplot$residuals))+
  geom_boxplot(data = boxplot$source, coef = params$boxplot_coefficient, outlier.colour = "red")+
  scale_x_discrete(breaks = NULL)+
  labs(x = NULL, y = boxplot$residuals)

if(nrow(boxplot$outliers) > 0){
  g <- g + geom_text(data = boxplot$outliers, aes(label = ID), size = 4, position = position_jitter(width = 0.2))
}

g
```
